Variation in use of East Asian Late Paleolithic weapons: A study of tip cross-sectional area of stemmed points from Korea

### 2022-10-03 analysis/paper/paper.qmd
Added text for the first results

## TCSA Variation

@@

ggplot(tcsa,
        aes(0.3, 
            min_tcsa)) +
  # show regions for tools from Table 3 of Lomabrd 2021
  # https://www-sciencedirect-com.offcampus.lib.washington.edu/science/article/pii/S0277379121002286
    annotate("rect", 
           xmin = 0.1,
           xmax = 0.2,
           ymin = 4,
           ymax = 18,
           fill = "green",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Poisoned\narrowheads",
               hjust = 0,
           x = point_function_group_label_y,
           y = (18-4) /2 + 4) +
      annotate("rect", 
           xmin = 0.2,
           xmax = 0.3,
           ymin = 13,
           ymax = 53,
           fill = "blue",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Arrowheads",
               hjust = 0,
           x = point_function_group_label_y,
           y = (53-13) /2 + 13) +
      annotate("rect", 
           xmin = 0.3,
           xmax = 0.4,
           ymin = 40,
           ymax = 76,
           fill = "orange",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Dart tips",
               hjust = 0,
           x = point_function_group_label_y,
           y = (76-40) /2 + 40 ) +
      annotate("rect", 
           xmin = 0.4,
           xmax = 0.5,
           ymin = 44,
           ymax = 98,
           fill = "purple",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Javelin",
               hjust = 0,
           x = point_function_group_label_y,
           y = (98-44) /2 + 44 ) +
     annotate("rect", 
           xmin = 0.5,
           xmax = 0.6,
           ymin = 88,
           ymax = 230,
           fill = "yellow",
           alpha = 0.2) +
      annotate("text", 
               label = "Thrusting\nspear",
               size = point_function_group_label_size,
               hjust = 0,
           x = point_function_group_label_y,
           y = (230-88) /2 + 88 ) +
  geom_boxplot(alpha = 0.1) +
  geom_quasirandom(alpha = 0.4,
                   size = 2.3) +
  coord_cartesian(xlim = c(-0.1,1)) +
  theme_bw(base_size = 6) +
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()
        ) +
  labs(y = "TCSA")
```
1) The mean TCSA of the Korean stemmed point is `r round(mean(tcsa$min_tcsa), digits=1)`, and standard deviation of TCSA is `r round(sd(tcsa$min_tcsa), digits=1)`. Overall we see wide variation of TCSA values in this boxplot.   
2) According to Lombard's (2021) TCSA ranges, Korean stemmed points are mostly in the categories of javelins and thrusting spear tips, with smaller numbers as dart tips and arrowheads.
3) Poison arrowheads seem to be absent from these Korean assemblages

>>>>

## TCSA range of Korean stemmed points

@@

ggplot(tcsa,
        aes(0.3, 
            min_tcsa)) +
  # show regions for tools from Table 3 of Lomabrd 2021
  # https://www-sciencedirect-com.offcampus.lib.washington.edu/science/article/pii/S0277379121002286
    annotate("rect", 
           xmin = 0.1,
           xmax = 0.2,
           ymin = 4,
           ymax = 18,
           fill = "green",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Poisoned\narrowheads",
               hjust = 0,
           x = point_function_group_label_y,
           y = (18-4) /2 + 4) +
      annotate("rect", 
           xmin = 0.2,
           xmax = 0.3,
           ymin = 13,
           ymax = 53,
           fill = "blue",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Arrowheads",
               hjust = 0,
           x = point_function_group_label_y,
           y = (53-13) /2 + 13) +
      annotate("rect", 
           xmin = 0.3,
           xmax = 0.4,
           ymin = 40,
           ymax = 76,
           fill = "orange",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Dart tips",
               hjust = 0,
           x = point_function_group_label_y,
           y = (76-40) /2 + 40 ) +
      annotate("rect", 
           xmin = 0.4,
           xmax = 0.5,
           ymin = 44,
           ymax = 98,
           fill = "purple",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Javelin",
               hjust = 0,
           x = point_function_group_label_y,
           y = (98-44) /2 + 44 ) +
     annotate("rect", 
           xmin = 0.5,
           xmax = 0.6,
           ymin = 88,
           ymax = 230,
           fill = "yellow",
           alpha = 0.2) +
      annotate("text", 
               label = "Thrusting\nspear",
               size = point_function_group_label_size,
               hjust = 0,
           x = point_function_group_label_y,
           y = (230-88) /2 + 88 ) +
  geom_boxplot(alpha = 0.1) +
  geom_quasirandom(alpha = 0.4,
                   size = 2.3) +
  coord_cartesian(xlim = c(-0.1,1)) +
  theme_bw(base_size = 6) +
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()
        ) +
  labs(y = "TCSA")
```

To gain a general idea of the role of stemmed points, we calculated TCSA for all stemmed points. @tcsa-all-sp shows the mean TCSA is `r round(mean(tcsa$min_tcsa), digits=1)`, and standard deviation of TCSA is `r round(sd(tcsa$min_tcsa), digits=1)`. Overall we see a wide variation of TCSA values in this boxplot. According to -@lombard2021variation TCSA ranges (Table 1), Korean stemmed points are mostly in the categories of javelins and thrusting spear tips, with smaller numbers as dart tips and arrowheads. Among the weapon-tip types, only poison arrowheads appear to be absent from these Korean assemblages.
